Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-26 ind_files/04_field_eqns.tex
Minor rephrasing

We will assume that our process is nice enough that we can find a function $P(\nu,t)$ such that the probability that the process takes value $\nu^{(t)}$ at time $t$ is given by\footnote{If you know some measure theory, note that this amounts to saying the following (I think?). Let $(\mathcal{M}(\mathcal{T}), \mathscr{F}, \{\mathscr{F}_t\}_{t\geq 0}, \mathbb{P})$ be the filtered probability space in which our adapted stochastic process $\{X_t\}_{t \geq 0}$ lives. Let $m$ be the Lebesgue measure on $(\mathcal{T},\mathcal{B}(\mathcal{T}))$. Then, we require $\mathbb{P}(X_t \ll m \ \forall \ t) = 1$, \emph{i.e.} we need the process to be $[\mathbb{P}]$-a.s. absolutely continuous with respect to the Lebesgue measure for all time. Since each $\nu \in \mathcal{M}(\mathcal{T})$ is a finite measure, it is obviously $\sigma$-finite, and we can now thus use the Radon-Nikodym theorem to find a density function $R_{m}(x,t)$ such that for any $t \geq 0$ and any $B \in \mathcal{B}(\mathcal{T})$, we have $X_t(B) = \int_B R_{m}(x,t) dm \ [\mathbb{P}]\textrm{-a.s.}$. We also require being able to find a second function $Q_{\mathbb{P}}(\nu,t)$ on $\mathscr{F} \times [0,\infty)$ such that we can write $\mathbb{P}(X_t = \nu) = \int_{\mathcal{T}} Q_{\mathbb{P}}(\nu,t)\nu(dx)$. The function $Q_{\mathbb{P}}(\nu,t)$ is again similar to a density function and thus essentially amounts to an absolute continuity condition, except on the stochastic process itself (and thus the measure $\mathbb{P}$ defined on $\mathscr{F}$) rather than on its state space (\emph{i.e.} the measures $\nu \in \mathcal{M}(\mathcal{T})$ defined on $\mathcal{T}$). We can now use the $[\mathbb{P}]$-a.s. assured absolute continuity with respect to the Lebesgue measure of any measure $\nu$ attained by our process $X_t$ to write $\int_{\mathcal{T}}Q_{\mathbb{P}}(\nu,t)\nu(dx) = \int_{\mathcal{T}}Q_{\mathbb{P}}(\nu,t)R_{m}(x,t)dm$. Changing notation from $dm$ to $dx$ to revert to the usual abuse of notation for integration with respect to the Lebesgue measure, we can now define $P(\nu,t)dm = Q_{\mathbb{P}}(\nu,t)R_{m}(x,t)dm$. This allows us finally to write $\mathbb{P}(X_t = \nu) =  \int_{\mathcal{T}}P(\nu,t)dx$ as desired.} $\int_{\mathcal{T}}P(\nu,t)dx$. In one dimension, this is often automatically satisfied for biologically reasonable choices of the birth and death functionals~(\cite{dawson_stochastic_1975,walsh_introduction_1986,konno_stochastic_1988,reimers_one_1989,dawson_stochastic_2000}; Also see section 2.3 in~\cite{etheridge_introduction_2000} and proposition 3.1 in~\cite{champagnat_individual_2008}).

>>>>

We will assume that our process is nice enough that we can find a function $P(\nu,t)$ such that the probability that the process takes value $\nu^{(t)}$ at time $t$ is given by\footnote{If you know some measure theory, note that this amounts to saying the following (I think?). Let $(\mathcal{M}(\mathcal{T}), \mathscr{F}, \{\mathscr{F}_t\}_{t\geq 0}, \mathbb{P})$ be the filtered probability space in which our adapted stochastic process $\{X_t\}_{t \geq 0}$ lives. Let $m$ be the Lebesgue measure on $(\mathcal{T},\mathcal{B}(\mathcal{T}))$. Then, we require $\mathbb{P}(X_t \ll m \ \forall \ t) = 1$, \emph{i.e.} we need the process to be $[\mathbb{P}]$-a.s. absolutely continuous with respect to the Lebesgue measure for all time. Since each $\nu \in \mathcal{M}(\mathcal{T})$ is a finite measure, it is obviously $\sigma$-finite, and we can now thus use the Radon-Nikodym theorem to find a density function $R_{m}(x,t)$ such that for any $t \geq 0$ and any $B \in \mathcal{B}(\mathcal{T})$, we have $X_t(B) = \int_B R_{m}(x,t) dm \ [\mathbb{P}]\textrm{-a.s.}$. We also require being able to find a second function $Q_{\mathbb{P}}(\nu,t)$ on $\mathscr{F} \times [0,\infty)$ such that we can write $\mathbb{P}(X_t = \nu) = \int_{\mathcal{T}} Q_{\mathbb{P}}(\nu,t)\nu(dx)$. The function $Q_{\mathbb{P}}(\nu,t)$ is similar to a density function and thus essentially amounts to another absolute continuity condition, except on the stochastic process itself (and thus the measure $\mathbb{P}$ defined on $\mathscr{F}$) rather than on the elements of its state space (\emph{i.e.} the measures $\nu \in \mathcal{M}(\mathcal{T})$ defined on $\mathcal{T}$). We can now use the $[\mathbb{P}]$-a.s. assured absolute continuity with respect to the Lebesgue measure of any measure $\nu$ attained by our process $X_t$ to write $\int_{\mathcal{T}}Q_{\mathbb{P}}(\nu,t)\nu(dx) = \int_{\mathcal{T}}Q_{\mathbb{P}}(\nu,t)R_{m}(x,t)dm$. Changing notation from $dm$ to $dx$ to revert to the usual abuse of notation for integration with respect to the Lebesgue measure, we can now define $P(\nu,t)dx \coloneqq Q_{\mathbb{P}}(\nu,t)R_{m}(x,t)dx$. This allows us finally to write $\mathbb{P}(X_t = \nu) =  \int_{\mathcal{T}}P(\nu,t)dx$ as desired.} $\int_{\mathcal{T}}P(\nu,t)dx$. In one dimension, this is often automatically satisfied for biologically reasonable choices of the birth and death functionals~(\cite{dawson_stochastic_1975,walsh_introduction_1986,konno_stochastic_1988,reimers_one_1989,dawson_stochastic_2000}; Also see section 2.3 in~\cite{etheridge_introduction_2000} and proposition 3.1 in~\cite{champagnat_individual_2008}).
